### 2021-02-10 response-ind.tex
add response-ind.tex

\documentclass[12pt,a4paper]{amsart}
\setlength{\textwidth}{\paperwidth}\addtolength{\textwidth}{-2in}\calclayout
\usepackage{hyperref}
\usepackage{comment}
\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\title[Response]{\large Response to the referee's comments}
\begin{document}
	\maketitle	
	We are grateful to the referee for the very helpful comments.
	We have incorporated all the comments made by the referee.
	Here are the changes in response to the comments of the first referee.

	\begin{itemize}
		\item ``It is, however, a pity that the branching mechanisms they deal with are not general enough(Assumption 2).''

	\end{itemize}
\begin{enumerate}
	\item ``Page 3: give clearly definition of $z^{1+\beta}$.''

	\item ``Page 5: In Lemma 2.1, introduce the notation $\mathcal P^+$;''

	\item ``page 6, Lemma 2.4: The same as above: define $x^{1+\beta}$, when $x$ is negative.''

	\item ``page 8: change `$T_uf$' into `$T_uf_s$'.''

	\item ``page 8: In the first inequality below (2.9), the last part missed $t^{-1}$.''

\end{enumerate}
	Here are the changes in response to the comments of the second reviewer.
	\begin{enumerate}
		\item ``Page 4. line 1. I think that `$\alpha \widetilde \beta \leq \kappa_f b$' should be `$\alpha \tilde \beta > \kappa_f b$'. ''

		\item ``Page 4, line 3. Correspondingly, `$\alpha \tilde \beta > \kappa_f b$' should be `$\alpha \tilde \beta \leq \kappa_f b$'. ''

		\item ``Page 9, line 15. `$=$' should be `:='.''

		\item ``Page 10, line 16. The term `$\langle f, \phi_p\rangle_{\varphi}$' was missed.''

	\end{enumerate}
\end{document}