### 2021-02-11 response-ind.tex
update response-ind.tex

\documentclass[12pt,a4paper]{amsart}
\setlength{\textwidth}{\paperwidth}\addtolength{\textwidth}{-2in}\calclayout
\usepackage{hyperref}
\usepackage{comment}
\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\title[Response]{\large Response to the referees' comments}
\begin{document}
	\maketitle	
	We are grateful to the referees for the very helpful comments.
	We have incorporated all the comments made by the referees.
	Here are the changes in response to the comments of the first referee.

	\begin{itemize}
		\item ``It is, however, a pity that the branching mechanisms they deal with are not general enough(Assumption 2).''

		\emph{We agree. We addressed this in  Remark 1.3.}

	\end{itemize}
\begin{enumerate}
	\item ``Page 3: give clearly definition of $z^{1+\beta}$.''

	\emph{We added the definition of $z^{1+\beta}$ after equation (1.2).}

	\item ``Page 5: In Lemma 2.1, introduce the notation $\mathcal P^+$;''

	\emph{We replaced ``$h\in \mathcal P^+$'' with ``non-negative $f\in \mathcal P$'' in the Lemma 2.1.}

	\item ``page 6, Lemma 2.4: The same as above: define $x^{1+\beta}$, when $x$ is negative.''

	\emph{In this paper, 
	the non-integer power for negative numbers is not defined. 
	We changed our statement in Lemma 2.4 and added a proof.}

	\item ``page 8: change `$T_uf$' into `$T_uf_s$'.''

	\emph{Changed as suggested.}

	\item ``page 8: In the first inequality below (2.9), the last part missed $t^{-1}$.''

	\emph{This is a typo. We added $t^{-1}$ back.}

\end{enumerate}
	Here are the changes in response to the comments of the second referee.
	\begin{enumerate}
		\item ``Page 4. line 1. I think that `$\alpha \widetilde \beta \leq \kappa_f b$' should be `$\alpha \tilde \beta > \kappa_f b$'. ''

		\emph{We changed our statement there to resolve this ambiguity. See lines 9-11 on page 4.}

		\item ``Page 4, line 3. Correspondingly, `$\alpha \tilde \beta > \kappa_f b$' should be `$\alpha \tilde \beta \leq \kappa_f b$'. ''

		\emph{Similar to the previous one.}

		\item ``Page 9, line 15. `$=$' should be `:='.''

		\emph{Changed as suggested.}

		\item ``Page 10, line 16. The term `$\langle f, \phi_p\rangle_{\varphi}$' was missed.''

		\emph{This is a typo. We added it.}

	\end{enumerate}
\end{document}